---
title: "NCollin_607_TidyVerse_Extend"
author: "Noah Collin"
date: "11/23/2021"
output: html_document
---

## Original

Link: https://raw.githubusercontent.com/acatlin/FALL2021TIDYVERSE/main/Data607CreateTidyverseVignette.Rmd

So this is what was posted to github that I'll be extending:

---
title: "Examples of tidyverse"
author: "Tom Buonora and ..."
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tidyverse Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
   warning = FALSE, 
  message = FALSE,
  fig.width = 8,
  fig.height = 8
)
```

```{r}

library(tidyverse)         # ggplot2, dplyr, tidyr, readr, tibble, sringr and more

```


# Tidyverse Vignette

##   readr : read_csv()

<br>
read_csv is part of readr whereas read.csv is base R. Im not sure that read_csv is tidier than read.csv
<br>

```{r readr : read_csv}
CURR_PATH<-str_trim(getwd())
# to do : use the kaggle api
# https://www.kaggle.com/mysarahmadbhat/nyc-traffic-accidents

# nyc_traffic_accidents<-read.csv("NYC Accidents 2020.csv")

nyc_traffic_accidents<-read_csv("https://raw.githubusercontent.com/TheReallyBigApple/CunyAssignments/main/DATA607/NYC%20Accidents%202020.csv",show_col_types = FALSE)


```

##   readr : spec()

the spec function displays the schema for the data frame or tibble

```{r readr : spec}

# the readr spec method shows you the schema
spec(nyc_traffic_accidents)


# this is base R, it removes the spaces from the column names
colnames(nyc_traffic_accidents) <- make.names(colnames(nyc_traffic_accidents))

```

##   dplyr : group_by()    dplyr_summarise()     magrittr : %>%

magrittr provides several operators including the commonly used %>% which pipes results from one function to the next

```{r several}


nyc_traffic_accidents %>%
  group_by(BOROUGH) %>%
  summarise(observations = n(), killed = sum(NUMBER.OF.PERSONS.KILLED, na.rm = TRUE))


```

##   dplyr : select()

select can select certain columns
select can also omit certain columns. 
Below I omit a sequence of columns starting with LATTITUDE and ending with LOCATION

```{r dplyr select}

nyc_traffic_accidents<-nyc_traffic_accidents %>% 
    select(!(LATITUDE:LOCATION))

```

##   dplyr : muatate()   replace_na()

muate() can create or modify a column
replace_na() can replace all instances of NA with a specified value

```{r replace_na}

nyc_traffic_accidents<-nyc_traffic_accidents %>% dplyr::mutate(BOROUGH = replace_na(BOROUGH, "NYC"))

```


##   ggplot :  geom_histogram() et all.

while the other tidyverse packages infer data parsing, ggplot exists to *display* data
having said that, its robust functionality includes data shaping. The histogram is essentially a group_by()
as is scale_x_dates

```{r ggplot histogram}


ggplot(data = nyc_traffic_accidents, aes(x = CRASH.DATE)) +
geom_histogram() +
theme(axis.text.x=element_text(angle=60, hjust=1)) +
ggtitle("Accidents Per Day") +
scale_x_date(date_breaks = "weeks", date_labels = "%Y-%m-%d") +
xlab("Date") 



```

## Extention

by *Noah Collin*

Facet wrap the above plot by borough: (also get rid of "NYC" borough, just disregard that data that doesn't have an appropriate borough.)
```{r facet1}
ggplot(data = nyc_traffic_accidents %>% filter(BOROUGH != "NYC"), aes(x = CRASH.DATE, fill = BOROUGH)) +
geom_histogram() +
theme(axis.text.x=element_text(angle=80, hjust=1)) +
ggtitle("Accidents Per Day, comparing boroughs") +
scale_x_date(date_breaks = "months", date_labels = "%Y-%m-%d") +
xlab("Date") + 
  facet_wrap(~ BOROUGH, nrow=2)

```


Now the same as above but free y axes to show the pattern over time in each indivual borough, not relative to one another:

Facet wrap the above plot by borough: (also get rid of "NYC" borough, just disregard that data)
```{r facet2}
ggplot(data = nyc_traffic_accidents %>% filter(BOROUGH != "NYC"), aes(x = CRASH.DATE, fill = BOROUGH)) +
geom_histogram() +
theme(axis.text.x=element_text(angle=80, hjust=1)) +
ggtitle("Accidents Per Day, boroughs with their own Y-Axis") +
scale_x_date(date_breaks = "months", date_labels = "%Y-%m-%d") +
xlab("Date") + 
  facet_wrap(~ BOROUGH, scales = "free_y", nrow=2)

```

There does seem to be a pattern here.  Accidents per day seem to be lowest in April but highest in February in each borough.  The pattern seems less pronounced in Staten Island.  That might be because it's more suburban there or becuase it's just a smaller y axis than all the other graphs.


For my own curiosity, I wonder what time of day looks like for each month time of day
```{r facet2}
ggplot(data = nyc_traffic_accidents %>% filter(BOROUGH != "NYC"), aes(x = CRASH.TIME, fill = BOROUGH)) +
geom_histogram() +
theme(axis.text.x=element_text(angle=80, hjust=1)) +
ggtitle("Amount of accidents per time of day") +
#scale_x_date(date_breaks = "months", date_labels = "%Y-%m-%d") +
xlab("Time") + 
  facet_wrap(~ BOROUGH, scales = "free_y", nrow=2)
```

I expected to see accidents spike around 9 a.m. and 5.p.m.  But that doesn't seem to be the case.

